Decomposition-based Job-shop Scheduling with Constrained Clustering

### 2021-11-29 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
reread Sections 5 and 6

In this work, we introduce a method based on clustering to automatically decompose a JSP instance into several time windows that are small enough for optimization by an ASP solver.
General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises close objects wrt.\ some distance measure. 
In scheduling settings, however, clusters must satisfy additional constraints implied by the precedence relation between operations of a job. Standard constrained clustering algorithms are not readily applicable to scheduling scenarios either, as they are limited to disjointness constraints specifying objects that must not appear together in a cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}. 
Therefore, our clustering method implements a novel type of constraints that
\begin{enumerate*}[label=\emph{(\roman*)}]
  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or a previous cluster and
  \item ensure balancing between clusters according to the target number of operations per cluster.
\end{enumerate*}
Further contributions of our work can be summarized as follows:
\begin{itemize}
  \item Since a typical JSP instance describing jobs, their operations, and available machines does not provide sufficient information by itself for finding some promising decomposition by a clustering algorithm, we incorporate heuristic approaches like \textit{First-In-First-Out} and \textit{Machine Load} to extract features from their solutions.
  \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
  \item The evaluation of our approach on Taillard's and Demirkol's benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
\end{itemize}
%It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations. The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window. The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan. 
%More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\

@@

We utilized the above feature sets for partitioning the operations of JSP instances into $3$ clusters or time windows, respectively,
and then ran the multi-shot optimization approach introduced in~\cite{el2020job} for
$10$ minutes in total, i.e., $200$ seconds per time window,
on an
%\todo{Please check and update the machine and operating system setup.}
Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
Dell Latitude 5590 machine under Windows 10.
%
The resulting quality of schedules in terms of the achieved makespan for each instance is reported in Table~\ref{tab4} and Table~\ref{tab5},
where the first four rows include the greedy FIFO and MTWR search methods supplied by the environment in~\cite{tagesc21a}
as well as single-shot and multi-shot pure ASP optimization, the latter taking earliest starting times instead of 
clustering for the decomposition into $3$ time windows,
for reference.
%
Regarding these reference methods, FIFO and MTWR are more or less on par, with a slight advantage of FIFO on Taillard's instances and an edge for MTWR on Demirkol's instances, 
indicated
by the average makespans given in the last column of each table.
While combinatorial optimization by neither single- nor multi-shot ASP 
solving is able to keep step within the time limit and leads to greater averages over all instances, the multi-shot pure ASP optimization approach turns out to be comparably robust
on half of the instances from Demirkol's benchmark suite,
which points out the strong impact of instance patterns.
%
\begin{table}[bt]
  \begin{center}
    \caption{Solution quality comparison for different feature sets (Demirkol's instances).}
    \label{tab5}
    \addtolength{\tabcolsep}{0.9pt}
    \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|} \hline
%      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
      \textbf{Approach}  & DE31  & DE32 & DE33 & DE34 & DE35 & DE71 & DE72 & DE73 & DE74 & DE75 & AVG  \\ \hline

@@

Among our feature sets for clustering, in preparation of multi-shot optimization by ASP,
we do not identify any clear winner dominating over all instances.
However, the feature sets \textbf{F2}, \textbf{F6}, \textbf{F11}, \textbf{F12}, \textbf{F14}, and \textbf{F16}
yield significant makespan improvements relative to the other alternatives on particular instances.
This means that considering the features \textit{RPT}, \textit{EST}, \textit{ML}, or
\wtfifo, \wtmtwr, and \wtest in addition to starting times \stfifo, \stmtwr, and \stest by
greedy algorithms is beneficial, while the particular combination of features achieving best schedules varies.
Moreover, our clustering method for JSP instance decomposition outperforms (multi-shot) pure ASP optimization,
and specific feature sets are ahead of the greedy FIFO and MTWR algorithms on seven out of ten Taillard's instances as well as half of Demirkol's instances,
as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}
and Table~\ref{tab5}.
Notably, the feature set \textbf{F6} leads to the shortest average makespans for
our cluster-wise JSP solving approach on both kinds of instances, comes close to
FIFO on Taillard's and is most robust on Demirkol's instances, but the
virtual best clustering results still improve significantly.
% In Table~\ref{tab5}, the feature sets outperformed the greedy on only three out of ten instances. However, feature set \textbf{F6} provided the best obtained average. 
We thus conclude that the instance decomposition by clustering can enable iterative combinatorial optimization
to find better schedules than greedy algorithms in a short amount of time, where alternative feature sets deserve
attention and trying several of them in parallel seems advisable.

@@

Data mining approaches~\cite{ismail2012production} have been extensively proposed and studied to generate % so-called
dispatching rules~\cite{blackstone1982state} for scheduling problems.
Unlike combinatorial and stochastic optimization methods, dispatching rules do not take
complete schedules into account, but address the prioritization of operations for making
local allocation decisions by greedy scheduling algorithms.
In addition to Earliest Starting Time (EST), First-In-First-Out (FIFO), and
Most Total Work Remaining (MTWR), which we also use for obtaining starting times as features,
common dispatching rules include Last-In-First-Out (LIFO) and policies incorporating
Job Length (JL) in terms of the number of operations as well as further features
like those described in Section~\ref{sec:features}:
Operation Position (OP), Processing Time (PT), Remaining Processing Time (RPT), 
Time Length of a Job (TLJ), and Machine Load (ML).
While dispatching rules enable an efficient operation allocation,
even under real-time conditions, the quality of their decisions heavily depends on the
problem instances under consideration, where data mining techniques come in to
improve over tedious and ad hoc manual tuning.

The approach of~\cite{koonce2000using} incorporates data mining methods,
using OP, PT, RPT, and ML as features, for generating dispatching rules that approximate
operation priorities found in schedules optimized by means of a genetic algorithm.
On randomly generated test instances of comparably small size, i.e., $36$ operations
per instance, the generated dispatching rules were shown to outperform greedy search
with shortest PT as a rigid criterion for making allocation decisions.
Genetic algorithms are also utilized in~\cite{harrath2002genetic} to obtain
promising solutions for JSP instance and extract features to consider for the
construction of dispatching rules.
The specifically generated three-tiered policy comprises shortest PT, smallest JL, and longest RPT as
dispatching criteria in the order of significance.
Note that such a composite dispatching rule needs to reconstructed from scratch when
new features or problem instances are analyzed, while our constrained clustering method
can readily be applied to different instance or feature sets, respectively.

Beyond static scheduling, dynamic JSP~\cite{french82a} deals with events, such as 
the release of new jobs or sudden machine breakdowns, along with rescheduling or
updating dispatching rules, respectively.
In analogy to static approaches,
the processing method of~\cite{shahzad2010discovering} applies tabu search to
training instances and generates allocation policies approximating the optimized
complete schedules based on features like EST, PT, and RPT of operations.
Applied in the dynamic JSP setting, the dispatching rules obtained by means of data mining
were shown to yield better or comparable performance to rigid allocation strategies from the
literature.
A rescheduling approach combining Variable Neighborhood Search (VNS)
with K-Means clustering is presented in~\cite{adibi2014clustering},
where clusters are updated to reconfigure the VNS on each event.
The performance of this combined rescheduling strategy in the dynamic JSP setting turned out
to be better than pure VNS and greedy algorithms with the FIFO, LIFO, or
shortest PT allocation policy.

@@

Similarities of our cluster-wise JSP solving approach to the discussed methods include
that we apply (fixed) dispatching rules to extract the starting times of operations
in corresponding schedules as instance features.
Notably, the constrained clustering algorithm in Section~\ref{sec:method} can readily
be applied to arbitrary feature sets for decomposing the operations of JSP instances
into balanced time windows, i.e., our clustering method does not presuppose tuning to
specific instances or features under consideration.
The iterative combinatorial optimization by time windows can be viewed as a trade-off
between efficient dispatching rules, aiming to optimize local allocation decisions by
prioritizing operations, and the stochastic or combinatorial optimization of
complete schedules for problem instances.
While instance size and complexity can make the search for optimal solutions
with state-of-the-art solvers for ASP, Mixed Integer Programming, or Constraint Programming
prohibitive~\cite{zhang2010hybrid},
their real-world application areas in scheduling include, e.g.,
industrial printing~\cite{balduccini11a},
team-building~\cite{rigralmaliiile12a},
shift design~\cite{abgemuscwo15c},
course timetabling~\cite{bainkaokscsotawa18a},
lab resource allocation~\cite{frscel21a},
and
medical treatment planning~\cite{dogagrmamopo21a}.

>>>>

In this work, we introduce a method based on clustering to automatically decompose a JSP instance into several time windows that are small enough for optimization by an ASP solver.
General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises close objects wrt.\ some distance measure. 
In scheduling settings, however, clusters must satisfy additional constraints implied by the precedence relation between operations of a job. Standard constrained clustering algorithms are not readily applicable to scheduling scenarios either, as they are limited to disjointness constraints specifying objects that must not appear together in a cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}. 
Therefore, our clustering method implements a novel type of constraints that
\begin{enumerate*}[label=\emph{(\roman*)}]
  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or a previous cluster and
  \item ensure balancing between clusters according to the target number of operations per cluster.
\end{enumerate*}
Further contributions of our work can be summarized as follows:
\begin{itemize}
  \item Since a typical JSP instance describing jobs, their operations, and available machines does not provide sufficient information by itself for finding some promising decomposition by a clustering algorithm, we incorporate heuristic approaches like \textit{First-In-First-Out} and \textit{Machine Load} to extract features from their solutions.
  \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
  \item The evaluation of our approach on Taillard's and Demirkol's benchmark instances \cite{taillard1993benchmarks, demirkol1998benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a short solving time limit of $10$ minutes.
\end{itemize}
%It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations. The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window. The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan. 
%More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\

@@

We utilized the above feature sets for partitioning the operations of JSP instances into $3$ clusters or time windows, respectively,
and then ran the multi-shot optimization approach introduced in~\cite{el2020job} for
$10$ minutes in total, i.e., $200$ seconds per time window,
on an
%\todo{Please check and update the machine and operating system setup.}
Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
Dell Latitude 5590 machine under Windows 10.
%
The resulting quality of schedules in terms of the achieved makespan for each instance is reported in Table~\ref{tab4} and Table~\ref{tab5},
where the first four rows include the greedy FIFO and MTWR search methods supplied by the environment in~\cite{tagesc21a}
as well as single-shot and multi-shot pure ASP optimization, the latter taking earliest starting times instead of 
clustering for the decomposition into $3$ time windows,
for reference.
%
Regarding these reference methods, FIFO and MTWR are more or less on par, with a slight advantage of FIFO on Taillard's instances and an edge for MTWR on Demirkol's instances, 
indicated
by the average makespans given in the last column of each table.
While combinatorial optimization by neither single- nor multi-shot ASP 
solving is able to keep step within the time limit and leads to greater averages over all instances, the multi-shot pure ASP optimization approach turns out to be comparably robust
on half of the instances from Demirkol's benchmark suite,
which points out the critical impact of instance patterns.
%
\begin{table}[bt]
  \begin{center}
    \caption{Solution quality comparison for different feature sets (Demirkol's instances).}
    \label{tab5}
    \addtolength{\tabcolsep}{0.9pt}
    \begin{tabular}{|l|c|c|c|c|c|c|c|c|c|c|c|} \hline
%      & \multicolumn{11}{|c|}{\textbf{Instances}} \\ \cline{2-12}
      \textbf{Approach}  & DE31  & DE32 & DE33 & DE34 & DE35 & DE71 & DE72 & DE73 & DE74 & DE75 & AVG  \\ \hline

@@

Among our feature sets for clustering, in preparation of multi-shot optimization by ASP,
we do not identify any clear winner dominating over all instances.
However, the feature sets \textbf{F2}, \textbf{F6}, \textbf{F11}, \textbf{F12}, \textbf{F13}, \textbf{F14}, and \textbf{F16}
yield significant makespan improvements relative to the other alternatives on particular instances.
This means that considering the features \textit{RPT}, \textit{EST}, \textit{ML}, or
\wtfifo, \wtmtwr, and \wtest in addition to starting times \stfifo, \stmtwr, and \stest by
greedy algorithms is beneficial, while the particular combination of features achieving best schedules varies.
Moreover, our clustering method for JSP instance decomposition outperforms (multi-shot) pure ASP optimization,
and specific feature sets are ahead of the greedy FIFO and MTWR algorithms on seven out of ten Taillard's instances as well as half of Demirkol's instances,
as indicated by the virtual best clustering results in the last row of Table~\ref{tab4}
and Table~\ref{tab5}.
Notably, the feature set \textbf{F6} leads to the shortest average makespans for
our cluster-wise JSP solving approach on both kinds of instances, comes close to
FIFO on Taillard's and is most robust on Demirkol's instances, but the
virtual best clustering results still improve significantly.
% In Table~\ref{tab5}, the feature sets outperformed the greedy on only three out of ten instances. However, feature set \textbf{F6} provided the best obtained average. 
We thus conclude that the instance decomposition by clustering can empower iterative combinatorial optimization
to find better schedules than greedy algorithms in a short amount of time, where alternative feature sets deserve
attention and trying several of them in parallel seems advisable.

@@

Data mining approaches~\cite{ismail2012production} have been extensively proposed and applied to generate % so-called
dispatching rules~\cite{blackstone1982state} for scheduling problems.
Unlike combinatorial and stochastic optimization methods, dispatching rules do not take
complete schedules into account, but address the prioritization of operations for making
local allocation decisions by greedy scheduling algorithms.
In addition to Earliest Starting Time (EST), First-In-First-Out (FIFO), and
Most Total Work Remaining (MTWR), which we also use for obtaining starting times as features,
common dispatching rules include Last-In-First-Out (LIFO) and policies incorporating
Job Length (JL) in terms of the number of operations as well as further features
like those described in Section~\ref{sec:features}:
Operation Position (OP), Processing Time (PT), Remaining Processing Time (RPT), 
Time Length of a Job (TLJ), and Machine Load (ML).
While dispatching rules enable an efficient operation allocation,
even under real-time conditions, the quality of their decisions heavily depends on the
problem instances under consideration, where data mining techniques come in to
improve over tedious and ad hoc manual tuning.

The approach of~\cite{koonce2000using} incorporates data mining methods,
using OP, PT, RPT, and ML as features, for generating dispatching rules that approximate
operation priorities found in schedules optimized by means of a genetic algorithm.
On randomly generated test instances of comparably small size, i.e., $36$ operations
per instance, the generated dispatching rules were shown to outperform greedy search
with shortest PT as a rigid criterion for making allocation decisions.
Genetic algorithms are also utilized in~\cite{harrath2002genetic} to obtain
promising solutions for JSP instances and extract features to consider for the
construction of dispatching rules.
The specifically generated three-tiered policy comprises shortest PT, smallest JL, and longest RPT as
dispatching criteria in the order of significance.
Note that such a composite dispatching rule needs to reconstructed from scratch when
new features or problem instances are analyzed, while our constrained clustering method
can readily be applied to different instance or feature sets, respectively.

Beyond static scheduling, dynamic JSP~\cite{french82a} deals with events, such as 
the release of new jobs or sudden machine breakdowns, along with rescheduling or
updating dispatching rules, respectively.
In analogy to the static setting,
the processing approach of~\cite{shahzad2010discovering} applies tabu search to
training instances and generates allocation policies approximating the optimized
complete schedules based on features like EST, PT, and RPT of operations.
Applied in the dynamic JSP setting, the dispatching rules obtained by means of data mining
were shown to yield better or comparable performance to rigid allocation strategies from the
literature.
A rescheduling approach combining Variable Neighborhood Search (VNS)
with K-Means clustering is presented in~\cite{adibi2014clustering},
where clusters are updated to reconfigure the VNS on each event.
The performance of this combined rescheduling strategy in the dynamic JSP setting turned out
to be better than pure VNS and greedy algorithms with the FIFO, LIFO, or
shortest PT allocation policy.

@@

Similarities of our cluster-wise JSP solving approach to the discussed methods include
that we apply (fixed) dispatching rules to extract the starting times of operations
in corresponding schedules as instance features.
Notably, the constrained clustering algorithm in Section~\ref{sec:method} can readily
be applied to arbitrary feature sets for decomposing the operations of JSP instances
into balanced time windows, i.e., our clustering method does not presuppose tuning to
specific instance or feature sets under consideration.
The iterative combinatorial optimization by time windows can be viewed as a trade-off
between efficient dispatching rules, aiming to optimize local allocation decisions by
prioritizing operations, and the stochastic or combinatorial optimization of
complete schedules for problem instances.
While instance size and complexity can make the search for optimal solutions
with state-of-the-art solvers for ASP, Mixed Integer Programming, or Constraint Programming
prohibitive~\cite{coltep19a,zhang2010hybrid},
successful % their real-world 
application areas in scheduling include, e.g.,
industrial printing~\cite{balduccini11a},
team-building~\cite{rigralmaliiile12a},
shift design~\cite{abgemuscwo15c},
course timetabling~\cite{bainkaokscsotawa18a},
lab resource allocation~\cite{frscel21a},
and
medical treatment planning~\cite{dogagrmamopo21a}.
